---
title: "Assignment 1"
author: "Rohit Jyotiba Chougule & 19200240"
date: "10/10/2019"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```


## Task 1: Manipulation

1. Load the dataset EurostatCrime2017.csv. Notice that the first column of the csv file contains the names of the countries that must be read as row names [Hint: Load in the file using the function read.csv]

```{r Task1Q1}
# Reading the csv file from the question to the variable- df
df <- read.csv("EurostatCrime2017.csv", row.names = 1)

```


2. What is the size (number of rows and columns) and the structure of this dataset?

```{r Task1Q2}
# Getting the dimensions of the data as number of rows and columns respectively:
dim(df)

# Structure of the dataset is as seen below:
str(df, strict.width="wrap")

#strict.width=wrap is useful here to wrap around the text in the output for the str()

```

The number of rows are 41 and columns are 12 as seen from the output for the code chunk Task1Q2

3. Produce appropriate commands to answer the following questions:
(i) For some countries Theft includes also burglary, and theft of motorised land vehicle, in others they are recorded separately. Add a new column called All Theft which contains the sum of all the crimes that have a theft component:
• Theft,
• Theft of a motorized land vehicle,
• Burglary,
• Burglary of private residential premises
Please consider NA values as 0 in this case. [Hint: you may want to use the function apply or rowSums, check their help files to see how to deal with missing values.] [1]
(ii) Remove the columns: Theft, Theft of a motorized land vehicle, Burglary, and Burglary of private residential premises.

```{r Task3Q1}
# Adding the column All_Totals to perform the sum of the columns:
All_Totals = rowSums(df[8:11], na.rm = TRUE)
df$All_Thefts <- All_Totals
All_Totals
```

(ii)
```{r Task3Q2}
# Removing the columns as suggested in the question:
df <- df[-c(8:11)]
colnames(df)
```


4. Work with the dataset you created in question 3ii, and list the countries that contain any missing data

```{r Task1Q4}
# Identifying all the rows with NA values in the dataset
test<-df[!complete.cases(df),]

#Printing the list of countries having NA values
test[1]
```

5. Remove the countries with missing data from the dataframe.

```{r Task1Q5}
# Usage of omit function to remove the NA values rows from the data set
new_df <- na.omit(df, invert=FALSE)

# Changing the row numbers to arrange the row numbers in a sequence 1 to n
rownames(new_df) <- 1:nrow(new_df)
```


6. How many observations and variables are in this new dataframe?
```{r Task1Q6}
# Dimension of the new dataset after having removed NA values
dim(new_df)

```

The number of rows are 25 and columns are 9 in the new data set.



## Task 2: Analysis

Work with the dataset produced in question 6. Produce appropriate commands to answer the following questions:

1. According to these data what were the 3 most common crimes in Ireland in 2017?

```{r Task2Q1}
# Identifying country Ireland from the entire dataset for further processsing
ireland_df <- new_df[which(new_df$Country=="Ireland"),]

# Assigning sequence to the rows for new subset data frame
#rownames(ireland_df) <- 1:nrow(ireland_df)

#ireland_df is the row or a small version of the data frame 
# which has records for the crimes in Ireland

# Taking the Column values of the subset data frame into a Vector for sorting
test <- unlist((ireland_df))

# Identifying the position of crimes in descending order.
pos <- order(test, decreasing = TRUE )

# Defining another vector crime_ire to append the actual crime 
# values for top3 crimes in Ireland

crime_ire <- c()

# Loop to append the values from Crime values in the crime_ire 

for(i in 1:3){
 crime_ire[i] <- c(test[pos[i]])
}

#typeof(ireland_df)

# Searching for the crime name for top three crimes from 
# crime_ire to print their names from the data frame


# Given below is the list of top 3 crimes in Ireland in 2017
for(i in 1:3){
  print((ireland_df[which(ireland_df==crime_ire[i])]))
}

```
All type of theft, Unlawful acts involving drugs and precursors, and Assault are the top 3 crimes in Ireland


2. Which country has the highest overall record of offences (per hundred thousand inhabitants)?
```{r Task2Q2}
Total_Offence = rowSums(new_df[2:9], na.rm = TRUE)

new_df$TotalCrime <- Total_Offence

Max_Crime = max(Total_Offence)

Country <- subset(new_df, TotalCrime==Max_Crime)

# The country with highest overall record of offences is:
Country[1]
```

Switzerland is the country with highest overall offence.


## Task 3: Creativity

Do something interesting with these data! Create a table or a plot which shows something we have not discovered above already and outline your findings.

```{r Task3}
# Taking the sum of all the columns except the column names to find the total crime 
# Taking a transpose of the sum
Total_Crime <- t(colSums(new_df[2:9]))

# Taking the data into the dataframe to get it into a readable format
All_Crime_df <- as.data.frame((Total_Crime))

# The below Table/dataframe shows most number of crimes in all the dataset
print(All_Crime_df, row.names = FALSE)

# Thefts is the crime which occurs most across all the countries from the given dataset
print(All_Crime_df[which.max(All_Crime_df)], row.names = FALSE)

# Intentional homicide is the least occured crime in all the countries from the given dataset

print(All_Crime_df[which.min(All_Crime_df)], row.names = FALSE)
```

```{r Task3BarPlot, fig.align='center'}
# The barplot as well shows the maximum number of crime overall
barplot(Total_Crime, beside = TRUE, las =2)

```

